---
title: "Transforming Non-Normal Data"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r child = "../setup.Rmd"}
```

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(gridExtra)
options(scipen = 999)
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

class: middle

## Review

Quiz  
Homework

---

```{r echo = FALSE, message = FALSE}
library(tidyverse)

library(broom)
theme_set(theme_gray(base_size = 18))
```


```{r}
library(AmesHousing)
ames_raw <- janitor::clean_names(ames_raw)
glimpse(ames_raw)
```

---

```{r fig.align='center', echo=FALSE, message=FALSE}
ggplot(ames_raw, aes(x = year_built, y = sale_price)) +
  geom_point() +
  scale_y_continuous(labels = scales::comma_format()) 
  
```


---


```{r fig.align='center', echo=FALSE}
ggplot(ames_raw, aes(x = year_built, y = log(sale_price))) +
  geom_point() +
  scale_y_continuous(labels = scales::comma_format()) 
  
```

Note that log is natural log in R.

---

```{r}
model_y <- lm(log(sale_price) ~ year_built, 
              data = ames_raw)
tidy(model_y)
```

${log(\hat y_i)} = b_0 + b_1x_{1i}$

${log(\hat y_i)} = -4.33 + 0.00829x_{1i}$
---

class: middle

Estimated sale price of a house built in 1980

${log(\hat y_i)} = -4.33 + 0.00829 \times 1980$

--

$e^{log(\hat y_i)} = e^{-4.33 + 0.00829 \times 1980}$

--

$\hat y_i = e^{-4.33} \times e^ {0.00829 \times 1980} = 177052.2$

--

For one-unit (year) increase in x, the y is multiplied by $e^{b_1}$.

--

Note that $e^x$ can be achieved by `exp(x)` in R.

---

class: middle

## Common Transformations

- $log(Y)$, $log(X)$
- $1/Y$, $1/X$
- $\sqrt{Y}, \sqrt{X})$
---

class: middle

Lecture activity in the weekly repo