---
title: "Model Evaluation and Validation"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r child = "../setup.Rmd"}
```

```{r echo = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(openintro)
library(broom)
library(modelr)
theme_set(theme_gray(base_size = 18))
options(scipen = 999)
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

class: middle

## Review

Quiz  
Homework

---


```{r}
glimpse(babies)
```

---

```{r}
model_g <- lm(bwt ~ gestation, data = babies)
tidy(model_g)
```

---

```{r}
glance(model_g)
```

---

```{r}
glance(model_g)$r.squared
```

16% of the variation in birth weight is explained by gestation. Higher values of R squared is preferred.

```{r echo = FALSE, fig.align='center', message=FALSE, warning = FALSE, fig.height = 4}
babies %>% 
  ggplot() +
  aes(x = gestation, y = bwt) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 
    
```

---

```{r echo = FALSE, fig.align='center', message=FALSE, warning = FALSE, fig.height = 4}
babies %>% 
  ggplot() +
  aes(x = gestation, y = bwt) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 
    
```

---

```{r}
model_gsa <- lm(bwt ~ gestation + smoke + age, data = babies)
```


---

## Adjusted R-Squared

.pull-left[
```{r}
glance(model_g)
```
]

.pull-right[
```{r}
glance(model_gsa)
```
]


---

```{r}
babies %>% 
  add_predictions(model_g) %>% 
  add_residuals(model_g) %>% 
  select(bwt, pred, resid)
```


---

```{r}
babies %>% 
  add_predictions(model_gsa) %>% 
  add_residuals(model_gsa) %>% 
  select(bwt, pred, resid)
```

---

## Root Mean Square Error

$RMSE = \sqrt{\frac{\Sigma_{i=1}^n(y_i-\hat y_i)^2}{n}}$
---

```{r}
modelr::rmse(model_gsa, babies)
```

```{r}
modelr::rmse(model_g, babies)
```

---

```{r}
model_full <- lm(bwt ~ gestation + parity + age +
                   height + weight + smoke, 
                 data = babies)
```

```{r}
modelr::rmse(model_full, babies)
```

```{r}
glance(model_full)$r.squared
```

Can we keep adding all the variables and try to get an EXCELLENT model fit?

---

## Overfitting

- We are fitting the model to sample data.

- Our goal is to understand the population data.

- If we make our model too perfect for our sample data, the model may not perform as well with other sample data from the population.

- In this case we would be overfitting the data.

- We can use **model validation** techniques.

---

## Splitting the Data (Train vs. Test)

```{r message=FALSE}
set.seed(12345)
babies_split <- rsample::initial_split(babies) ## 75% to 25% split
```

--

```{r}
babies_train <- rsample::training(babies_split)
dim(babies_train)
```

--

```{r}
babies_test <- rsample::testing(babies_split)
dim(babies_test)
```

---

```{r}
model_gsa_train <- lm(bwt ~ gestation + smoke + age, data = babies_train)
model_gsa_test <- lm(bwt ~ gestation + smoke + age, data = babies_test)
```

---

.pull-left[
```{r}
glance(model_gsa_train)
```
]

.pull-right[
```{r}
glance(model_gsa_test)
```
]


---

.pull-left[
```{r}
modelr::rmse(model_gsa_train, babies_train)
```
]

.pull-right[
```{r}
modelr::rmse(model_gsa_test, babies_test)
```
]


---

class: middle inverse

.font50[Evaluation of Logistic Regression Models]

---

class: middle

```{r}
babies <- babies %>% 
  mutate(low_bwt = case_when(bwt < 88 ~ TRUE,
                             bwt >= 88~ FALSE)) %>% 
  drop_na(gestation)
```

---
class: middle

```{r}
model_g <- glm(low_bwt ~ gestation, 
               data = babies,
               family = "binomial")
```

---
class: middle

```{r}
tidy(model_g)
```

$\hat p = \frac{\exp(b_0 +b_1x)}{1+\exp(b_0 + b_1x)}$

$\hat p$ when gestation is 284 = $\frac{\exp(17.5 -0.0758 \cdot 284)}{1+\exp(17.5 -0.0758 \cdot 284)} = \frac{\exp(-4.0272)}{1+ \exp(-4.0272)} = 0.01751203$

---

```{r}
babies <- babies %>% 
  add_predictions(model_g) 
select(babies, pred) 
```

---

```{r}
babies <- babies %>% 
  mutate(pred_p = exp(pred)/(1+exp(pred)))
select(babies, pred, pred_p)
```

---

## Cutoff 

```{r}
babies <- babies %>% 
  mutate(pred_y = case_when(pred_p < 0.5 ~ FALSE, 
                           pred_p >= 0.5 ~ TRUE))
select(babies, low_bwt, pred, pred_p, pred_y)
```

---

### Confusion Matrix

```{r}
janitor::tabyl(babies, low_bwt, pred_y) %>% 
  janitor::adorn_totals(c("row", "col"))
```

**Sensitivity** (true-positive rate): 4/57 = 0.0702  
**Specificity** (true-negative rate): 1161/1166 = 0.9957  
**Accuracy**: (1161+4)/1223 = 0.9526